Supporting Information for Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-09-07 si.tex
fixing ht and fc integrals

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}
In this section, we discuss the evaluation of Herzberg-Teller integrals, in the limit of no Duschinsky coupling, when the ground and first excited state are harmonic wells.
We use the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation}
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates.
Franck-Condon factors and Herzberg-Teller integrals are often evaluated in position space using known solutions to the Harmonic Oscillator Hamiltonian \cite{RN230}
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N(m) H_m(q) \exp(-q^2/2)
\end{equation}
where $N(m) = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the m$^{th}$ Hermite Polynomial.
We begin by generating an expression for $\mel{n}{q^k}{m}$, as this will resolve into Franck-Condon factors for $n = 0$ and Herzberg-Teller integrals for $n=1$.

By letting $\ket{m}$ belong to the ground state surface $|g)$ and $\ket{n}$ belong to the first excited surface $|e)$, we write
\begin{equation}
\begin{split}
		\mel{n}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \psi^*_n(q-\Delta) q^k \psi_m(q)\\
		&= \int_{-\infty}^\infty \mathrm{d}q N(n) N(m) H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k H_m(q) \exp(-\frac{q^2}{2})
\end{split}
\end{equation}
Introducing $\tilde{N} \equiv N(n)N(m)$ and slightly rearranging, we see
\begin{equation}\label{key}
	\mel{n}{q^k}{m} = \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q H_n(q+\Delta) q^k H_m(q) \exp(-q^2 + q\Delta)
\end{equation}
This is the key result.
As a side note, \autoref{key} explicitly shows
\begin{subequations}
	\begin{equation}
		\lim_{\Delta \rightarrow 0} \langle n | m \rangle = \delta_{n,m}
	\end{equation}
and
	\begin{equation}
	\lim_{\Delta \rightarrow 0} \mel{n}{q}{m} = \sqrt{\frac{1}{2}} \left(\sqrt{m}\delta_{n,m-1} + \sqrt{m+1} \delta_{n,m+1}\right)
	\end{equation}
\end{subequations}
i.e., as the offset vanishes, the normal harmonic oscillator selection rules are obtained, the expected result, where $\delta$ is the Kronecher delta.

@@

We now have the background needed to calculate $\mel{1}{q}{0}$ for an arbitrary offset $\Delta$.
We note that Myers et al. have developed a general expression to evaluate Franck-Condon factors for two displaced harmonic wells of arbitrary curvature,\cite{Myers1982} so we focus our attention onto Herzberg-Teller integrals, which are not commonly tabulated.
Starting with \autoref{key},
\begin{equation}
\begin{split}
	\mel{1}{q}{0} = \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q H_1(q+\Delta) q H_0(q) \exp(-q^2 - q\Delta)
\end{split}
\end{equation}
Using $H_0(q) = 1$ and $H_1(q) = 2q$,\cite{RN230} 
\begin{equation}\label{HT10}
	\begin{split}
		\mel{1}{q}{0} &= \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ 2(q + \Delta) q \exp(-q^2 - q\Delta)\\
		&= 2\tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ (q^2 + q \Delta) \exp(-q^2 - q\Delta)\\
		&=  2\tilde{N} \exp(-\frac{\Delta^2}{2}) \left(\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 - q\Delta) + 2\Delta \int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 - q\Delta) \right)
	\end{split}
\end{equation}
These integrals are now easily evaluated using \autoref{identity}:
\begin{subequations}\label{evalGaussians}
	\begin{equation}
		\begin{split}
					\int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 - q\Delta) &= -\partial_{\Delta} \left(\sqrt{\pi} \exp(\frac{\Delta^2}{4})\right)\\
					& = -\sqrt{\pi} \frac{\Delta}{2} \exp(\frac{\Delta^2}{4})
		\end{split}
	\end{equation}
giving
\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 - q\Delta) &= -\partial_{\Delta} \left(-\sqrt{\pi} \left(\frac{\Delta}{2} \exp(\frac{\Delta^2}{4})\right)\right)\\
		& =\sqrt{\pi} \left(\frac{1}{2}\exp(\frac{\Delta^2}{4}) + \frac{\Delta^2}{4} \exp(\frac{\Delta^2}{4})\right)
	\end{split}
\end{equation}
\end{subequations}
With $\tilde{N} = N(1)*N(0) = (1! \ 2^1 \pi^{\frac{1}{2}})^{-\frac{1}{2}} * (0! \ 2^0 \pi^{\frac{1}{2}})^{-\frac{1}{2}} = \sqrt{1/2\pi}$ and substituting the results of \autoref{evalGaussians} into \autoref{HT10}, 
\begin{equation}
	\mel{1}{q}{0} = \sqrt{\frac{1}{2}} (1 - \frac{\Delta^2}{2}) \exp(\frac{-\Delta^2}{4})
\end{equation}
The evaluation of other Herzberg-Teller integrals follows a similar process, or by using symbolic manipulation packages. 
The relevant Franck-Condon factors and Herzberg-Teller integrals, assuming identical curvature and no Duschinsky coupling, are tabulated below.

>>>>

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}
In this section, we discuss the evaluation of Herzberg-Teller integrals, in the limit of no Duschinsky coupling, when the ground and first excited state are harmonic wells.
We use the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation}
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates.
Franck-Condon factors and Herzberg-Teller integrals are often evaluated in position space using known solutions to the Harmonic Oscillator Hamiltonian \cite{RN230}
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N(m) H_m(q) \exp(-q^2/2)
\end{equation}
where $N(m) = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the m$^{th}$ Hermite Polynomial.
We begin by generating an expression for $\mel{\tilde{n}}{q^k}{m}$, as this will resolve into Franck-Condon factors for $n = 0$ and Herzberg-Teller integrals for $n=1$.
As in the main text, $\{\ket{\tilde{a}}\}$ are vibrational states on the first excited state $|e)$ and $\{\ket{b}\}$ are vibrational states on the ground state $|g)$.

By letting $\ket{m}$ belong to the ground state surface $|g)$ and $\ket{\tilde{n}}$ belong to the first excited surface $|e)$, we write
\begin{equation}
\begin{split}
		\mel{\tilde{n}}{q^k}{m} &= \int_{-\infty}^\infty \mathrm{d}q \psi^*_n(q-\Delta) q^k \psi_m(q)\\
		&= \int_{-\infty}^\infty \mathrm{d}q N(n) N(m) H_n(q-\Delta) \exp(-\frac{(q - \Delta)^2}{2}) q^k H_m(q) \exp(-\frac{q^2}{2})
\end{split}
\end{equation}
Introducing $\tilde{N} \equiv N(n)N(m)$ and slightly rearranging, we see
\begin{equation}\label{key}
	\mel{\tilde{n}}{q^k}{m} = \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q H_n(q+\Delta) q^k H_m(q) \exp(-q^2 + q\Delta)
\end{equation}
This is the key result.
As a side note, \autoref{key} explicitly shows
\begin{subequations}
	\begin{equation}
		\lim_{\Delta \rightarrow 0} \langle \tilde{n} | m \rangle = \langle n | m \rangle = \delta_{n,m}
	\end{equation}
and
	\begin{equation}
	\lim_{\Delta \rightarrow 0} \mel{\tilde{n}}{q}{m} = \mel{{n}}{q}{m} = \sqrt{\frac{1}{2}} \left(\sqrt{m}\delta_{n,m-1} + \sqrt{m+1} \delta_{n,m+1}\right)
	\end{equation}
\end{subequations}
i.e., as the offset vanishes, the normal harmonic oscillator selection rules are obtained, the expected result, where $\delta$ is the Kronecher delta.

@@

We now have the background needed to calculate $\mel{\tilde{1}}{q}{0}$ for an arbitrary offset $\Delta$.
We note that Myers et al. have developed a general expression to evaluate Franck-Condon factors for two displaced harmonic wells of arbitrary curvature,\cite{Myers1982} so we focus our attention onto Herzberg-Teller integrals, which are not commonly tabulated.
Starting with \autoref{key},
\begin{equation}
\begin{split}
	\mel{\tilde{1}}{q}{0} = \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q H_1(q+\Delta) q H_0(q) \exp(-q^2 - q\Delta)
\end{split}
\end{equation}
Using $H_0(q) = 1$ and $H_1(q) = 2q$,\cite{RN230} 
\begin{equation}\label{HT10}
	\begin{split}
		\mel{\tilde{1}}{q}{0} &= \tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ 2(q + \Delta) q \exp(-q^2 - q\Delta)\\
		&= 2\tilde{N} \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ (q^2 + q \Delta) \exp(-q^2 - q\Delta)\\
		&=  2\tilde{N} \exp(-\frac{\Delta^2}{2}) \left(\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 - q\Delta) + 2\Delta \int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 - q\Delta) \right)
	\end{split}
\end{equation}
These integrals are now easily evaluated using \autoref{identity}:
\begin{subequations}\label{evalGaussians}
	\begin{equation}
		\begin{split}
					\int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 - q\Delta) &= -\partial_{\Delta} \left(\sqrt{\pi} \exp(\frac{\Delta^2}{4})\right)\\
					& = -\sqrt{\pi} \frac{\Delta}{2} \exp(\frac{\Delta^2}{4})
		\end{split}
	\end{equation}
giving
\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 - q\Delta) &= -\partial_{\Delta} \left(-\sqrt{\pi} \left(\frac{\Delta}{2} \exp(\frac{\Delta^2}{4})\right)\right)\\
		& =\sqrt{\pi} \left(\frac{1}{2}\exp(\frac{\Delta^2}{4}) + \frac{\Delta^2}{4} \exp(\frac{\Delta^2}{4})\right)
	\end{split}
\end{equation}
\end{subequations}
With $\tilde{N} = N(1)*N(0) = (1! \ 2^1 \pi^{\frac{1}{2}})^{-\frac{1}{2}} * (0! \ 2^0 \pi^{\frac{1}{2}})^{-\frac{1}{2}} = \sqrt{1/2\pi}$ and substituting the results of \autoref{evalGaussians} into \autoref{HT10}, 
\begin{equation}
	\mel{\tilde{1}}{q}{0} = \sqrt{\frac{1}{2}} (1 - \frac{\Delta^2}{2}) \exp(\frac{-\Delta^2}{4})
\end{equation}
The evaluation of other Herzberg-Teller integrals follows a similar process. 
The relevant Franck-Condon factors and Herzberg-Teller integrals, assuming identical curvature and no Duschinsky coupling, are tabulated below.
